Windowed least-squares model reduction for dynamical systems

### 2019-10-07 section3/space_time.tex
latest

\subsection{\spaceTimeAcronym\ trial spaces}
We now briefly consider \spaceTimeAcronym\ trial subspaces. 
\KTC{Error in first equation. This is the approximate solution not the real
one. Use parallel language to hwo this was done in the last section for
S-reduction trial subspaces. Also make sure the basis is zero at the beginning
of the interval and the affine transformation is taken to be the solution at
the end of the last window. This will not be too trivial, as the solution on
the previous slab enters the trial subspace definition.}
Over 
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\KTC{Next sentence, use parallel language to the last section. It's subtly
different but I think written better there.} Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
the \spaceTimeAcronym\ trial subspace computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptArg{n} ), \\ 
      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =
  \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}\stgenstateArg{n-1})(\timeEndArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMArgnt{1})(0)
 & n=1. \end{cases}
%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}
\KTC{In prior equation, remove projection and align properly with ST-LSPG
stuff before.}

>>>>

\subsection{\spaceTimeAcronym\ trial spaces}
\KTC{Error in first equation. This is the approximate solution not the real
one. Use parallel language to hwo this was done in the last section for
S-reduction trial subspaces. Also make sure the basis is zero at the beginning
of the interval and the affine transformation is taken to be the solution at
the end of the last window. This will not be too trivial, as the solution on
the previous slab enters the trial subspace definition.}

We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptArg{n}\otimes \onesFunctionArg{n} .
\end{equation}
Here $\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}} \otimes \timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function}. At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

\end{comment}
\KTC{Next sentence, use parallel language to the last section. It's subtly
different but I think written better there.} Substituting the approximation~\eqref{eq:stapprox} into the minimization problem~\eqref{eq:tclsrm}, \methodAcronym\ with 
the \spaceTimeAcronym\ trial subspace computes the sequence of generalized coordinates $\stgenstateArg{n}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,
\begin{align}\label{eq:obj_gen_slab_spacetime}
\begin{split}
&\underset{\stgenstateyArg{n} \in \RR{\stdim}}{\text{minimize}}\; \mathcal{J}^n( \stbasisArgnt{n} \stgenstateyArg{n} + \stateInterceptArg{n} ), \\ 
      &\text{subject to }\;  \stbasisArg{n}{\timeStartArg{n}}\stgenstateyArg{n}  =
  \begin{cases} 
\mathbb{P}^n(\stbasisArgnt{n-1}(\timeEndArg{n-1})\stgenstateArg{n-1})  & n = 2,\ldots,\nslabs, \\ 
\mathbb{P}^n(\stateFOMIC)
 & n=1. \end{cases}
%\begin{cases} \genstateyArg{n-1}{\timeEndArg{n-1}} & n = 2,\ldots,\nslabs \\
%\genstateIC & n=1, \end{cases} 
\end{split}
\end{align}
\KTC{In prior equation, remove projection and align properly with ST-LSPG
stuff before.}

### 2019-10-07 section4/numerical_methods.tex
latest

\section{Numerical solution techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
As \methodAcronym\ displays commonalities with optimal control problems, a large existing body of work can be leveraged. This section discuses a variety of such techniques and outlines how they can be used to solve the \methodAcronym\ system. We focus primarily on direct and indirect methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%Equation~\ref{eq:obj_gen_slab} displays commonalities with optimal control problems. For timeslabs $n = 1,\ldots,\nslabs$, the objective is to find the \textit{function} $\genstate^n(t)$ that minimizes the \textit{functional} $ \mathcal{J}^n$. The distinction between Eq.~\ref{eq:obj_gen_slab} and traditional optimal control problems is that, in the form presented in Eq.~\ref{eq:obj_gen_slab}, (1) there is not a controller and (2) there is no system that describes the dynamics for the generalized coordinates. 
%Techniques used to solve optimal control problems can be leveraged to solve Eq.~\ref{eq:obj_gen_slab} (or, equivalently, Eqs.~\eqref{eq:tclspg_oc1} and~\eqref{eq:tclspg_oc2}).

\subsection{\spatialAcronym: direct and indirect methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Direct method eschew the calculus of variations 
and attempt to ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order optimality conditions~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". For both direct and indirect methods, a variety of
discretization/solution approaches are then possible. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

In the present context, both direct and indirect methods are investigated to solve \methodAcronym\ with \spatialAcronym\ trial subspaces. We consider:

\begin{itemize} \item Discretize then Optimize (Direct Methods): 
\begin{itemize}
\item Discretize and then optimize approaches are outlined for linear multistep methods. 
\end{itemize} 
\item Optimize then Discretize (Indirect Methods): 
\begin{itemize} 
\item An indirect method leveraging the forward-backward sweep algorithm is outlined 
to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
 \end{itemize} 
\end{itemize} 
It
is emphasized that a variety of other approaches exist. A detailed exploration
of additional approaches will be a subject of future work. Section~\ref{sec:direct} 
will outline direct solution approaches based on linear multistep methods, 
while Section~\ref{sec:indirect} will outline the indirect solution approach.

>>>>

\section{Numerical solution techniques}\label{sec:numerical_techniques}
The discussion to this point has outlined the formulation of the \methodAcronym\ method. We now consider numerical solution techniques. 
\EP{I think passive voice is fine in this instance, it reads easier}
As \methodAcronym\ displays commonalities with optimal
control problems, a large existing body of work can be leveraged. This section
discuses a variety of such techniques and outlines how they can be used to
solve the \methodAcronym\ system. We focus primarily on 
%\KTC{earlier we kept
%saying 
%direct (i.e.,
%	discretize-then-optimize) and indirect (i.e., optimize-then-discretize).
%	Let's be consistent when we re-introduce it here.}
direct (discretize-the-optimize) and indirect (optimize-then-discretize) for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces.

\subsection{\spatialAcronym: direct and indirect methods}
Numerical techniques to
solve optimal control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}.  Direct methods eschew the first order optimality conditions
~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} and, instead, ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system and objective functional and ``transcribing"
the infinite dimensional problem into a finite-dimensional nonlinear
programming (optimization) problem. Direct approaches thus ``discretize then optimize."
Indirect methods, on the other hand, leverage the calculus of variations to
derive the Euler--Lagrange equations, which comprise the first-order
optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} associated with the objective 
functional. Indirect methods then solve (numerically) the Euler--Lagrange equations. Thus, indirect methods ``optimize then discretize," and 
solve the optimization problem
``indirectly". A variety of discretization/solution techniques are possible for both direct and indirect methods. Collocation methods,
finite elements, spectral methods, shooting methods, etc., are all examples of
plausible solution techniques.

In the present context, we investigate both direct and indirect methods to solve \methodAcronym\ with \spatialAcronym\ trial
subspaces. We consider: 
\begin{itemize} \item \textit{Direct Methods (discretize-then-optimize)}: Discretize and then optimize approaches are outlined for linear multistep methods. 
\item \textit{Indirect methods (optimize-then-discretize)}: An indirect method leveraging the forward-backward sweep algorithm is outlined 
to solve~\eqref{eq:lspg_continuous}-\eqref{eq:lspg_bcs}. 
\end{itemize} 
We emphasize that a variety of other approaches exist. Future work will comprise a detailed exploration
of additional approaches.
Section~\ref{sec:direct}  outlines direct solution approaches based on linear
multistep methods, while Section~\ref{sec:indirect} outlines the indirect
solution approach.
